\LARGE \bf Eye tracking with Educational Robots: A Cautionary Tale

### 2014-09-30 paper.tex
with previous commit
Removed old figures + temporary files
Updated figures (to a Latex style)
Updated author list (still hidden for review)
While here, hide the acknowledgment section as well
[formatting] Typeset conditions in sans serif

%\numberofauthors{1} 
%\author{
%\alignauthor
%Julia Fink, S\'{e}verin Lemaignan, Pierre Dillenbourg\\
%%\titlenote{Dr.~Trovato insisted his name be first.}
%       \affaddr{Computer-Human Interaction in Learning and Instruction Lab (CHILI)}\\
%       \affaddr{Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL)}\\
%       \affaddr{CH-1015 Lausanne}\\
%       \affaddr{Switzerland}\\
%       \email{firstname.lastname@epfl.ch}
%}
%
%\additionalauthors{Additional authors: 
%Francesco Mondada, LSRO, EPFL, francesco.mondada@epfl.ch 
%}
%

@@

The experiment consists in two phases. The first phase is the observation phase
during which the participants watch the robot move and detect an obstacle in a
``playground''. The second phase requires the participants to interact with the
robot by placing an obstacle in its path to make it detect the obstacle before
reaching a certain target.  After each phase the participant explain the
behavior of the robot. The sensor data was visualized through LEDs with three
visualization conditions: TRUE visualization, RANDOM visualization and NONE
visualization. Through this experiment we investigate the following research
questions:

@@

In the experiment, there were 3 sensor data visualization conditions.  For the
TRUE visualization condition, the number of illuminated LEDs, on the top of the
robot, is proportional to the intensity measured by the IR-sensor. When the
intensity of the reflected IR crosses a predefined threshold, the robot does a
90° turn to its right. For the two other conditions, the behavior is the same,
simply the display changes. In the RANDOM condition, the display shows as if the
robot was detecting the obstacle at random moments, while in the NONE
visualization condition nothing is shown on top of the robot. The robot has a
fitting for a pencil that we used for the drawing of its trajectory.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{expe}
    \caption{\small Experimental setup for observation phase. The observation
    video is recorded from the video camer in the front of the eye tracker (in
    the lop-left corner).}

@@

We recruited 52 participants for the experiment. All of them were
students from École Polytechnique Fédérale de Lausanne, Switzerland.
Most of them were in their first year undergraduate program (with two
exceptions). ~The age distribution was narrow, with a mean of 19.7 and a
standard deviation of 1.97. There were 13 female and 39 male
participants. In the three experimental conditions, NONE, RANDOM and
TRUE, there were 18, 17 and 17 participants respectively.

@@

The average fixation duration on the robot during observation phase
(Figure~\ref{res1}) is significantly more in TRUE and RANDOM condition than in
the NONE condition ($F[2,49]=3.68$, p = .03). This depicts that the
sensor data visualization has an effect on the users' attention.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\linewidth]{res1}
    \caption{Average fixation duration on the robot ~during the observation phase
    vs.~Condition}
    \label{res1}
\end{figure}

@@

The first improvement is significantly more in NONE and RANDOM
conditions (Figure~\ref{res2}) than in the TRUE condition ($F[2,49]=3.75$, p =
0.03).

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\linewidth]{res2}
    \caption{First improvement vs.~experimental condition}
    \label{res2}
\end{figure}

@@

The average fixation time on the reference side of the playground
(Figure~\ref{res3}) during the interaction phase is significantly more in NONE
and RANDOM condition than it is in the TRUE condition ($F[2,49]=4.19$,
p = .02).

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\linewidth]{res3}
    \caption{Average fixation duration on the reference side during the
    interaction phase vs.~Condition}
    \label{res3}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\linewidth]{res4}
    \caption{First improvement in centimeters ($y$-axis) vs.~number of fixations
    on reference side ($x$-axis) during the interaction phase.}
    \label{res4}
\end{figure}

@@

We present an eye-tracking study in the context of human-robot
interaction with an educational setting. The fact that participants have
higher average fixation duration on the robot during the observation
phase in the NONE condition than the other two conditions (Figure 4) is
not surprising as displaying visual information on the robot induces a
saliency effect on the attention and the gaze is attracted towards the
salient feature in the field of view.

The fact that the participants in the RANDOM or NONE visualization
condition improved more than those with TRUE condition is surprising
(Figure 5). There are two plausible explanations. The first is that
participants who see the robot in the TRUE condition have a stronger
believe that the robot always behaves the same. They see the LEDs of the
robot turning on and therefore the robot still works. The ones with NONE
visualization do not have an indication whether the robot works the same
way as in the observation phase, and start experimenting earlier. The
second is that the participants in TRUE condition just put the obstacles
out of detection range, because they could see different display
information and concluded therefore it may behave differently.

The fact that participants have lower average fixation duration on the
reference side in the interaction phase in the TRUE condition than the
other two conditions actually supports the explanation that participants
with TRUE condition had a higher initial belief in their cognitive model
of how does the robot work (Figure 6). They did not need to look onto
the reference given in observation phase. This also verifies our working
hypothesis that a carefully designed mapping between the robot's
behavior and its visual actuators can help in maximizing the learning
effect during the interaction.

@@

Another plausible explanation for the participants in NONE and RANDOM condition
being better performers than the TRUE condition during the interaction phase
comes from the transparency of the robot's behavior~\cite{??}. The robot's
behavior was not at all transparent in the former two conditions. This might
have caused the participants to put more effort in understanding the behavior.
Therefore, they performed better.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Acknowledgments}

This research was supported by...

>>>>

%\numberofauthors{1} 
%\author{
%\alignauthor
%Kshitij Sharma, S\'{e}verin Lemaignan, Pierre Dillenbourg\\
%       \affaddr{Computer-Human Interaction in Learning and Instruction Lab (CHILI)}\\
%       \affaddr{Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL)}\\
%       \affaddr{CH-1015 Lausanne}\\
%       \affaddr{Switzerland}\\
%       \email{firstname.lastname@epfl.ch}
%}
%
%\additionalauthors{Additional authors: 
%Francesco Mondada, LSRO, EPFL, francesco.mondada@epfl.ch 
%}
%

@@

The experiment consists in two phases. The first phase is the observation phase
during which the participants watch the robot move and detect an obstacle in a
``playground''. The second phase requires the participants to interact with the
robot by placing an obstacle in its path to make it detect the obstacle before
reaching a certain target.  After each phase the participant explain the
behavior of the robot. The sensor data was visualized through LEDs with three
visualization conditions: {\sf TRUE} visualization, {\sf RANDOM} visualization and {\sf NONE}
visualization. Through this experiment we investigate the following research
questions:

@@

In the experiment, there were 3 sensor data visualization conditions.  For the
{\sf TRUE} visualization condition, the number of illuminated LEDs, on the top of the
robot, is proportional to the intensity measured by the IR-sensor. When the
intensity of the reflected IR crosses a predefined threshold, the robot does a
90° turn to its right. For the two other conditions, the behavior is the same,
simply the display changes. In the {\sf RANDOM} condition, the display shows as if the
robot was detecting the obstacle at random moments, while in the {\sf NONE}
visualization condition nothing is shown on top of the robot. The robot has a
fitting for a pencil that we used for the drawing of its trajectory.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{setup}
    \caption{\small Experimental setup for observation phase. The observation
    video is recorded from the video camer in the front of the eye tracker (in
    the lop-left corner).}

@@

We recruited 52 participants for the experiment. All of them were
students from École Polytechnique Fédérale de Lausanne, Switzerland.
Most of them were in their first year undergraduate program (with two
exceptions). ~The age distribution was narrow, with a mean of 19.7 and a
standard deviation of 1.97. There were 13 female and 39 male
participants. In the three experimental conditions, {\sf NONE}, {\sf RANDOM} and
{\sf TRUE}, there were 18, 17 and 17 participants respectively.

@@

The average fixation duration on the robot during observation phase
(Figure~\ref{res1}) is significantly more in {\sf TRUE} and {\sf RANDOM} condition than in
the {\sf NONE} condition ($F[2,49]=3.68$, p = .03). This depicts that the
sensor data visualization has an effect on the users' attention.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\linewidth]{meanPlotFixRobo}
    \caption{Average fixation duration on the robot ~during the observation phase
    vs.~Condition}
    \label{res1}
\end{figure}

@@

The first improvement is significantly more in {\sf NONE} and {\sf RANDOM}
conditions (Figure~\ref{res2}) than in the {\sf TRUE} condition ($F[2,49]=3.75$, p =
0.03).

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\linewidth]{meanPlotFirstImprove}
    \caption{First improvement vs.~experimental condition}
    \label{res2}
\end{figure}

@@

The average fixation time on the reference side of the playground
(Figure~\ref{res3}) during the interaction phase is significantly more in {\sf NONE}
and {\sf RANDOM} condition than it is in the {\sf TRUE} condition ($F[2,49]=4.19$,
p = .02).

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\linewidth]{meanPlotFixReference}
    \caption{Average fixation duration on the reference side during the
    interaction phase vs.~Condition}
    \label{res3}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\linewidth]{corPlotFirstImprove}
    \caption{First improvement in centimeters ($y$-axis) vs.~number of fixations
    on reference side ($x$-axis) during the interaction phase.}
    \label{res4}
\end{figure}

@@

We present an eye-tracking study in the context of human-robot
interaction with an educational setting. The fact that participants have
higher average fixation duration on the robot during the observation
phase in the {\sf NONE} condition than the other two conditions (Figure 4) is
not surprising as displaying visual information on the robot induces a
saliency effect on the attention and the gaze is attracted towards the
salient feature in the field of view.

The fact that the participants in the {\sf RANDOM} or {\sf NONE} visualization
condition improved more than those with {\sf TRUE} condition is surprising
(Figure 5). There are two plausible explanations. The first is that
participants who see the robot in the {\sf TRUE} condition have a stronger
believe that the robot always behaves the same. They see the LEDs of the
robot turning on and therefore the robot still works. The ones with {\sf NONE}
visualization do not have an indication whether the robot works the same
way as in the observation phase, and start experimenting earlier. The
second is that the participants in {\sf TRUE} condition just put the obstacles
out of detection range, because they could see different display
information and concluded therefore it may behave differently.

The fact that participants have lower average fixation duration on the
reference side in the interaction phase in the {\sf TRUE} condition than the
other two conditions actually supports the explanation that participants
with {\sf TRUE} condition had a higher initial belief in their cognitive model
of how does the robot work (Figure 6). They did not need to look onto
the reference given in observation phase. This also verifies our working
hypothesis that a carefully designed mapping between the robot's
behavior and its visual actuators can help in maximizing the learning
effect during the interaction.

@@

Another plausible explanation for the participants in {\sf NONE} and {\sf RANDOM} condition
being better performers than the {\sf TRUE} condition during the interaction phase
comes from the transparency of the robot's behavior. The robot's
behavior was not at all transparent in the former two conditions. This might
have caused the participants to put more effort in understanding the behavior.
Therefore, they performed better.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Removed for double-blind review
%\section*{Acknowledgments}
%
%This research was supported by...
